### 2020-10-16 conclusion.tex
Final version.

\subsubsection*{Conclusion and Future Work}
We presented Priority GV, a~session-typed functional language which uses priorities to ensure deadlock freedom, and showed its relation to Priority CP~\cite{dardhagay18} via an operational correspondence.

>>>>

\subsubsection*{Conclusion and future work}
We presented Priority GV, a~session-typed functional language which uses priorities to ensure deadlock freedom, and showed its relation to Priority CP~\cite{dardhagay18} via an operational correspondence.

### 2020-10-16 priority-gv.tex
Final version.

The only non-standard feature of the typing rules is the priority annotations. Priorities are based on \emph{obligations/capabilities} used by Kobayashi~\cite{kobayashi06}, and simplified to single priorities following Padovani~\cite{padovani14}. The integration of priorities into GV is adapted from Padovani and Novara~\cite{padovaninovara15}. Paraphrasing Dardha and Gay~\cite{dardhagay18}, priorities obey the following two laws:
\begin{enumerate}
  \item an action with priority $\cs{o}$ happens before an action with priority $\cs{o+n}$;
  \item communication requires \emph{equal} priorities for the dual actions.
\end{enumerate}

@@

While our configuration typing is based on the standard typing for GV, it differs on two counts
$(i)$ we require that child threads return the unit value, as opposed to a terminated channel;
$(ii)$ we simplify typing for parallel composition.
In order to guarantee deadlock freedom, in GV each parallel composition must split \emph{exactly one} channel of the channel pseudo-type $\ty{S^\sharp}$ into two endpoints of type $\ty{S}$ and $\ty{\co{S}}$. Consequently, associativity of parallel composition does not preserve typing. In PGV, we guarantee deadlock freedom using priorities, which removes the channel pseudo-type $\ty{S^\sharp}$, simplifies typing for parallel composition, and restores type preservation for the structural congruence.

@@

\begin{compacttheorems}
  \begin{definition}[Canonical Forms]
    \label{def:pgv-canonical-forms}
    A~configuration $\tm{\conf{C}}$ is in canonical form if it is of the following form, where no term $\tm{M_i}$ is a value:
    \[
      \tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\main\;N)}}}.
    \]
  \end{definition}
  \begin{definition}[Normal Forms]
    A~configuration $\tm{\conf{C}}$ is in normal form if it is of the following form, where each $\tm{M_i}$ is ready to act on $\tm{x_i}$:
    \[
      \tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\main\;V)}}}.
    \]
  \end{definition}
  \begin{restatabletheorem}{thmpgvclosedprogressconfs}[Progress, $\cred$]
    \label{thm:pgv-closed-progress-confs}
    If $\cseq[\main]{\emptyenv}{\conf{C}}$ and $\tm{\conf{C}}$ is in canonical form, then either $\tm{\conf{C}}\cred\tm{\conf{D}}$ for some $\tm{\conf{D}}$; or $\tm{\conf{C}\equiv\conf{D}}$ for some $\tm{\conf{D}}$ in normal form.
  \end{restatabletheorem}
  \begin{proof}
    Our proof follows the reasoning by Kobayashi~\cite{kobayashi06} used in the proof of deadlock freedom for closed processes (Theorem 2). Details are in~\Cref{prf:thm-pgv-closed-progress-confs}.
  \end{proof}
  \begin{corollary}
    \label{cor:pgv-closed-progress}
    If $\cseq[\main]{\emptyenv}{\conf{C}}$, $\tm{\conf{C}}\centernot\cred$, and $\tm{\conf{C}}$ contains no endpoints, then $\tm{\conf{C}}\equiv\tm{\phi\;V}$ for some value V.
  \end{corollary}
\end{compacttheorems}
\Cref{thm:pgv-closed-progress-confs} and \cref{cor:pgv-closed-progress} state that a configuration in PGV will complete all its communication actions, thus satisfying deadlock freedom or reduce to a term which is a value in PGV and contains no channel endpoints.
\end{document}

>>>>

The only non-standard feature of the typing rules is the priority annotations. Priorities are based on \emph{obligations/capabilities} used by Kobayashi~\cite{kobayashi06}, and simplified to single priorities following Padovani~\cite{padovani14}. The integration of priorities into GV is adapted from Padovani and Novara~\cite{padovaninovara15}. Paraphrasing Dardha and Gay~\cite{dardhagay18}, priorities obey the following two laws:
\begin{enumerate*}[label=(\roman*)]
\item an action with priority $\cs{o}$ happens before an action with priority $\cs{o+n}$; and
\item communication requires \emph{equal} priorities for the dual actions.
\end{enumerate*}

@@

While our configuration typing is based on the standard typing for GV, it differs on two counts:
\begin{enumerate*}[label=(\roman*)]
\item we require that child threads return the unit value, as opposed to a terminated channel; and
\item we simplify typing for parallel composition.
\end{enumerate*}
In order to guarantee deadlock freedom, in GV each parallel composition must split \emph{exactly one} channel of the channel pseudo-type $\ty{S^\sharp}$ into two endpoints of type $\ty{S}$ and $\ty{\co{S}}$. Consequently, associativity of parallel composition does not preserve typing. In PGV, we guarantee deadlock freedom using priorities, which removes the channel pseudo-type $\ty{S^\sharp}$, simplifies typing for parallel composition, and restores type preservation for the structural congruence.

@@

\begin{compacttheorems}
  \begin{definition}[Canonical Forms]
    \label{def:pgv-canonical-forms}
    A~configuration $\tm{\conf{C}}$ is in canonical form if it is of the following form, where no term $\tm{M_i}$ is a value:
    \[
      \tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\main\;N)}}}.
    \]
  \end{definition}
  \begin{definition}[Normal Forms]
    A~configuration $\tm{\conf{C}}$ is in normal form if it is of the following form, where each $\tm{M_i}$ is ready to act on $\tm{x_i}$:
    \[
      \tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\main\;V)}}}.
    \]
  \end{definition}
  \begin{restatabletheorem}{thmpgvclosedprogressconfs}[Progress, $\cred$]
    \label{thm:pgv-closed-progress-confs}
    If $\cseq[\main]{\emptyenv}{\conf{C}}$ and $\tm{\conf{C}}$ is in canonical form, then either $\tm{\conf{C}}\cred\tm{\conf{D}}$ for some $\tm{\conf{D}}$; or $\tm{\conf{C}\equiv\conf{D}}$ for some $\tm{\conf{D}}$ in normal form.
  \end{restatabletheorem}
  \begin{proof}
    Our proof follows the reasoning by Kobayashi~\cite{kobayashi06} used in the proof of deadlock freedom for closed processes (Theorem 2). Details are in~\Cref{prf:thm-pgv-closed-progress-confs}.
  \end{proof}
  \begin{corollary}
    \label{cor:pgv-closed-progress}
    If $\cseq[\main]{\emptyenv}{\conf{C}}$, $\tm{\conf{C}}\centernot\cred$, and $\tm{\conf{C}}$ contains no endpoints, then $\tm{\conf{C}}\equiv\tm{\phi\;V}$ for some value V.
  \end{corollary}
\end{compacttheorems}
It follows from \cref{thm:pgv-closed-progress-confs} and \cref{cor:pgv-closed-progress} that a term which does not return an endpoint will complete all its communication actions, thus satisfying deadlock freedom.
\end{document}
